---
title: "SMPE: CO2 Concentration"
author: "Quentin Guilloteau"
output:
  html_document:
    number_sections: true
    df_print: paged
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objectives

There are two main objectives for this work:

1. Generate a graph showing a periodical oscillation superposed with a slower global evolution of the CO2

2. Caracterize the oscillation and propose a model of the global trend. We will also give an estimation of the CO2 concentration. 

# Importing the necessary Libraries

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Downloading the data

We first need to download the weekly measures from the [Scripps CO2 Program Website](https://scrippsco2.ucsd.edu):
```{r}
# Downloading the data
csv_url <- "https://scrippsco2.ucsd.edu/assets/data/atmospheric/stations/in_situ_co2/weekly/weekly_in_situ_co2_mlo.csv"
df <- read.csv(csv_url, sep = ",", comment.char = "\"", blank.lines.skip=TRUE, strip.white=TRUE, skip=44, header=FALSE)
```

We are skipping 44 lines because the ``csv`` file contains a header of 44 lines of comments.

```{r}
head(df)
```

The columns in the file are not named, so we call them ``date`` and ``CO2``:
```{r}
# Naming the columns
colnames(df) <- c("date", "CO2")
```

Let us take a quick look at the data and see the type of each column:
```{r}
# A quick look at the data
head(df)
sapply(df, typeof)
```

We see that the type of the column ``date`` is not ``Date`` but ``r typeof(df$date)``.

We will thus convert it to a type ``Date``.

```{r}
df$date <- as.Date(df$date)
sapply(df, class)
```

We will now check that there is no ``NA`` values in the data set.
```{r}
# Checking if there are NAs in the data
na_lines <- apply(df, 1, function(x) any(is.na(x)))
number_of_na_lines <- nrow(df[na_lines,])
number_of_na_lines
```
There are ``r number_of_na_lines`` ``NA`` in the dataset.

Let us remove the potential lines containing ``NA``:
```{r}
df <- df %>% drop_na()
```

# Discovering the Dataset

Let us discover the dataset and simply plot the evolution of the CO2 concentration from 1958 to present days:
```{r}
p <- ggplot(data = df, aes(x = date, y = CO2)) +
        theme_linedraw() +
        geom_point(size = 0.1) +
        xlab("Date") + ylab("CO2 Concentation (in ppm)") +
        ggtitle("Evolution of the CO2 Concentration from 1958 to present days")
p
```

The previous graph is zoomed in the data, but let us take a step back and look with the ``y`` axis starting at 0:


```{r}
p + expand_limits(y = 0)
```

We can see that there seems to be a periodical oscillation but also the global trend is increasing.

# Regression for global trend

## Setting up the Data

Instead of using pure dates, we will convert each date into the number of days since the first measurement of the data set
```{r}
date_first_measure <- df$date[1]
df$day <- as.integer(df$date - date_first_measure)
head(df)
```

## Linear Regression

We can try to fit a line for the CO2 concentration:

```{r}
# Linear Regression
reg_line <- lm(data = df, CO2 ~ day)
summary(reg_line)
```

We can plot the estimation of this model:

```{r}
# Compute the predictions
df$prediction_line <- predict(reg_line, df)
# Plot the data and the predictions
ggplot(data = df, aes(x = date, y = CO2)) +
    theme_linedraw() +
    geom_point(size = 0.1) +
    geom_line(aes(x = date, y = prediction_line), colour = "blue") +
    xlab("Date") + ylab("CO2 (in ppm)") +
    ggtitle("CO2 Variations and its Linear Estimation")
```

We can see that even if the $R^2$ value in the summary was close to 1, a line does not seem to be a good fit for this data.

Indeed, this model...

* ... is underestimating from year 1958 to 1970

* ... is overestimating from yaer 1970 to 2005

* ... is underestimating from year 2005 to the present days

## Quadratic Regression

As we saw previously, a line is not a good fit to model the data.

Let us try with a polynom of degree 2:
$$
CO2(day) = a \times day^2 + b \times day + c
$$

A polynom of degree 2 seems adapted to this situation because the evolution of the CO2 concentration looks slighlty quadratic.

We can use the function ``poly`` to quickly generate of polynom.

```{r}
reg_quad <- lm(data = df, CO2 ~ poly(day, 2))
summary(reg_quad)
```

We can now plot the data and the predictions:

```{r}
df$prediction_quad <- predict(reg_quad, df)
ggplot(data = df, aes(x = date, y = CO2)) +
    theme_linedraw() +
    geom_point(size = 0.1) +
    geom_line(aes(x = date, y = prediction_quad), colour = "blue") +
    xlab("Date") + ylab("CO2 (in ppm)") +
    ggtitle("CO2 Variations and its Quadratic Estimation")
```

The quadratic model fits way more the data.

## CO2 Prediction for 2025

As our regression takes the number of days from ``r date_first_measure``, we have to substract this date to have a valid entry.

We will do a prediction of the CO2 concentration evolution from the current date to the January 1st 2025.

```{r}
# January, 1st 2025
date_2025 <- as.Date("2025-01-01")
day_2025 <- as.integer(date_2025 - date_first_measure)

# Current date
date_today <- Sys.Date()
day_today <- as.integer(date_today - date_first_measure)
```

```{r}
# Creating points to estimate
data_prediction <- data.frame(day = seq(day_today, day_2025, 7))
data_prediction$date <- as.Date(date_first_measure + data_prediction$day)
# Predicting the CO2 Concentration
data_prediction$CO2 <- predict(reg_quad, data_prediction)

ggplot() +
    theme_linedraw() +
    geom_point(data = df, aes(x = date, y = CO2), size = 0.1) +
    geom_line(data = data_prediction, aes(x = date, y = CO2), color = "red") + 
    xlab("Date") + ylab("CO2 (in ppm)") +
    ggtitle("CO2 Variations to the Present Day and its Estimation until 2025")
```

We can say that if the CO2 concentration in the air continues to follow this quadratic grow, the concentration of CO2 in the air in 2025 will be around 425 ppm.

# Yearly Evolution

We will add a column in our data frame to indicate the which day of the year this is January 1st being day 0 and December 31st being 365.
This will help us compare each year.

```{r}
df$year <- as.integer(format(df$date, "%Y"))
df$day_of_year <- as.integer(df$date - as.Date(paste(df$year, "01", "01", sep = "-")))
```

We also want to look at the variations of CO2 in the year during the year.
So we will substract the value of the first measure of the year for every measure of this year.

```{r}
df$CO2_year <- 0
df$percentage_increase_year <- 0
for (y in df$year) {
  first_CO2_value_of_the_year <- df[df$year == y,]$CO2[1]
  df[df$year == y,]$CO2_year <- df[df$year == y,]$CO2 - first_CO2_value_of_the_year
  df[df$year == y,]$percentage_increase_year <- df[df$year == y,]$CO2 * 100 / first_CO2_value_of_the_year - 100
}
```

We can plot the yearly evolution of some years:

```{r}
ggplot(data = subset(df, year == 1960 | year == 1997 | year == 2019), 
       aes(x = day_of_year, y = percentage_increase_year, color = factor(year))) +
    theme_linedraw() +
    geom_line() +
    scale_y_continuous(limits = c(-2, 2)) +
    xlab("Day of the Year") + ylab("Percentage of CO2 Concentration increase") + labs(color = "Year") +
    ggtitle("CO2 Evolution for the Years 1960, 1997 and 2019")
```

We see that the concentration of CO2 is oscillating during the year with a maximal value reached around day 150 (month of May) and a minimal value around day 260 (month of September).


## Sinusoidal Regression

Let us try to fit a sinusoidal curve for the oscillations over one year.
We decide to add a polynomial expression to represent the global evolution over the year.

As we saw previously, the global trend of the CO2 concentration over the year is quadratic on the year.
We thus add a polynomial term of degree 2 to represent this evolution.

```{r}
reg_sin <- lm(data = df, percentage_increase_year ~ sin(2 * pi * day_of_year / 366) +
                                    cos(2 * pi * day_of_year / 366) +
                                    poly(day_of_year, 2) +
                                    poly(year, 2))
summary(reg_sin)
```

```{r}

df$prediction_sin <- predict(reg_sin, df)

ggplot(data = df[df$year == 1997,]) +
    theme_linedraw() +
    geom_point(aes(x = day_of_year, y = percentage_increase_year)) +
    geom_line(aes(x = day_of_year, y = prediction_sin)) +
    scale_y_continuous(limits = c(-2, 2)) +
    xlab("Days of the Year") + ylab("CO2 Variation (in ppm)") +
    ggtitle("CO2 Variations and its Sin Prediction for the Year 1997 (in ppm)")
```

## Polynomial Regression

Let us try with a polynomial regression to see if we can get a better model.

We have to chose a degree greater than 2 to get such an oscillation.
We decided to use a polynom of degree 5.

Once again we add a polynom of degree 2 to represent the global quadratic evolution of the CO2 concentration.

```{r}
degree <- 5
reg_poly <- lm(data = df, percentage_increase_year ~ poly(day_of_year, degree) + poly(year, 2))
summary(reg_poly)
```

```{r}
df$prediction_poly <- predict(reg_poly, df)

ggplot(data = df[df$year == 2019,]) +
    theme_linedraw() +
    geom_point(aes(x = day_of_year, y = percentage_increase_year)) +
    geom_line(aes(x = day_of_year, y = prediction_poly)) +
    scale_y_continuous(limits = c(-2, 2)) +
    xlab("Days of the Year") + ylab("CO2 Variation (in ppm)") +
    ggtitle("CO2 Variations and its Polynimial Prediction for the Year 1997 (in ppm)")
```

# Conclusion

We showed that the global evolution of the CO2 concentration was increasing overtime.
We also showed that there was a periodical oscillation during a time span of one year.

You can find the source code here [https://github.com/GuilloteauQ/SMPE](https://github.com/GuilloteauQ/SMPE).